---
title: "Conservation Funding Threats EDA"
author: "Jay Schoen, Michael Levin"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
library(scales)
library(ggsci)
```

# Data Import/Prep
```{r, echo = FALSE}
data <- read_csv("~/Columbia E3B/Conservation Group/Output V15.2/master_output_data_analysis_v15_filtered.csv") %>%
  as.data.frame() %>%
  dplyr::select(c(1:42, 49:109, 43:48))

data.multi <- read_csv("~/Columbia E3B/Conservation Group/Output V15.2/master_output_data_analysis_v15_mutlicategories.csv") %>%
  as.data.frame()

maxwell <- read_csv("~/Columbia E3B/Conservation Group/Maxwell et al table S2.csv") %>%
  as.data.frame()
```

## Category Lists
```{r, echo = FALSE}
clim.energy_cat <- names(data)[c(37,38,66:74,81)]         # climate/energy/combos/other

biodiv.bio.taxa.hab_cat <- names(data)[c(19:28,43, 30:36, 82:103)]
biodiv.bio.taxa.hab_cat_multi <- names(data.multi)[c(19:28,30:36)]

u.threat_cat <- names(data)[37:42]
u.taxa_cat <- names(data)[30:36]

all.threat_cat <- append(u.threat_cat, names(data)[66:81])
all.taxa_cat <- append(u.taxa_cat, names(data)[82:103])

u.threat_list <- c("Climate Change", 
                   "Energy & Transportation",
                   "Habitat Loss",
                   "Invasive Species",
                   "Overexploitation",
                   "Pollution")
u.taxa_list <- c("Amphibians",
                 "Birds",
                 "Fish",
                 "Invertebrates",
                 "Mammals",
                 "Plants",
                 "Reptiles")
```

## Colors
```{r, echo = FALSE}
cat_colors22 <- c("#990000", "#BF1E1E", "#993399", "#4C0099", "#000099",
                  "#2E10C3", "#3685B3", "#1AB3B3", "#8FC04F", "#09A256", 
                  "#006400", "#FFF928", "#C6C600", "#FFA200", "#CE8300", 
                  "#D0471E", "#9F8109", "#835506", "#317575", "#808080",
                  "black", "white")
barplot(1:22, col = cat_colors22, names.arg = 1:22)

cat_colors7 <- cat_colors22[c(1, 5, 20, 7, 18, 11, 19)]
barplot(1:7, col = cat_colors7)
```

## Grant Subsets
### No Keyword Grants
```{r}
no_keywords <- data %>% 
  filter(Keywords =="['-']") %>%
  arrange(desc(Grant.Amount))
no_keywords_plot <- ggplot(no_keywords[1:50,],
                            aes(x = factor(row.names(no_keywords[1:50,]), levels = 1:50),
                                y = Grant.Amount)) + 
  geom_col() +
  theme_minimal() +
  labs(title = "No Keyword Biggest Grants (50)",
       x = "Rank",
       y = "Money ($US)") +
  scale_y_continuous(labels = scales::comma)
no_keywords_plot
```

### All Other Grants
```{r}
data <- filter(data, Keywords != "['-']") %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002)
data.multi <- filter(data.multi, Keywords != "['-']") %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002)

money.data <- data[,11:(length(data)-6)]*data[,7]
money.data <- mutate(money.data, Year.Authorized = data$Year.Authorized)

by.year.money <- money.data %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_taxa_other), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))

```


### Bio Subset
Removing climate/energy grants that do not contain biodiversity conservation keywords
```{r}
# Single Category Data
data_bio.1 <- data %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat), any_vars(.> 0)) 
data_non.bio.1 <- data %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat), all_vars(. == 0))

data_bio.2 <- data_non.bio.1 %>%
  filter_at(vars(clim.energy_cat), all_vars(.==0))
data_non.bio.2 <- data_non.bio.1 %>%
  filter_at(vars(clim.energy_cat), any_vars(.>0))

data_bio <- bind_rows(data_bio.1, data_bio.2)
data_non.bio <- data_non.bio.2

# Multi-category Data
data.multi_bio.1 <- data.multi %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat_multi), any_vars(.> 0)) 
data.multi_non.bio.1 <- data.multi %>%
  filter_at(vars(biodiv.bio.taxa.hab_cat_multi), all_vars(. == 0))

data.multi_bio.2 <- data.multi_non.bio.1 %>%
  filter_at(vars(cat_threat_climate, cat_threat_energytransport),
            all_vars(.==0))
data.multi_non.bio.2 <- data.multi_non.bio.1 %>%
  filter_at(vars(cat_threat_climate, cat_threat_energytransport),
            any_vars(.>0))

data.multi_bio <- bind_rows(data.multi_bio.1, data.multi_bio.2)
data.multi_non.bio <- data.multi_non.bio.2
```

### Money Data
```{r}
# Single Category Data
money.data_bio <- data_bio[,11:(length(data_bio)-6)]*data_bio[,7]
money.data_bio <- mutate(money.data_bio,
                         Year.Authorized = data_bio$Year.Authorized,
                         Num.Thr = data_bio$Num.Thr)

by.year.money_bio <- money.data_bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_taxa_other), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Year.Authorized) %>%
  mutate(ann.percent = Amount/sum(Amount)) %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))

money.data_non.bio <- data_non.bio[,11:(length(data_non.bio)-6)]*data_non.bio[,7]
money.data_non.bio <- mutate(money.data_non.bio, 
                             Year.Authorized = data_non.bio$Year.Authorized,
                             Num.Thr = data_non.bio$Num.Thr)

by.year.money_non.bio <- money.data_non.bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_threat_overexp_pollution), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))

# Multi-category Data
money.data.multi_bio <- data.multi_bio[,11:(length(data.multi_bio)-6)]*data.multi_bio[,7]
money.data.multi_bio <- mutate(money.data.multi_bio, 
                               Year.Authorized = data.multi_bio$Year.Authorized,
                               Num.Thr = data.multi_bio$Num.Thr)

by.year.money.multi_bio <- money.data.multi_bio %>%
  group_by(Year.Authorized) %>%
  summarize_at(vars(cat_act_area:cat_threat_pollution), sum, na.rm = TRUE) %>%
  pivot_longer(-Year.Authorized, names_to = "Category", values_to = "Amount") %>%
  group_by(Category) %>%
  mutate(percent = Amount/sum(Amount))
```

# Figure 1: Proportion of Money vs. Species Threatened
```{r echo=FALSE}
fig.1.money_bio <- by.year.money_bio %>%
  filter(Category %in% u.threat_cat) %>%
  group_by(Category) %>%
  summarise(Total = sum(Amount))

fig.1.species <- maxwell %>% 
  dplyr::select(Category, n_species_threatened) %>%
  group_by(Category) %>%
  summarise(Total = sum(n_species_threatened)) 

rawFrame <- left_join(fig.1.money_bio, fig.1.species, by = "Category") %>%
  mutate(rawMoney = Total.x) %>%   # Total Money
  mutate(rawSpecies = Total.y) %>% # Total Species
  dplyr::select(Category, rawMoney, rawSpecies)

fig.1.data_bio <- left_join(fig.1.money_bio, fig.1.species, by = "Category") %>%
  mutate(Money = (Total.x/sum(Total.x))*100) %>%   # Percentage of Total Money
  mutate(Species = (Total.y/sum(Total.y))*100) %>% # Percentage of Total Species
  dplyr::select(Category, Money, Species) %>%
  pivot_longer(c(Money, Species), names_to = "is.Money", values_to = "Amount") %>%
  left_join(rawFrame, by = "Category") %>%
  mutate(rawMoney = ifelse(is.Money == "Species", NA, rawMoney)) %>%
  mutate(rawSpecies= ifelse(is.Money == "Money", NA, rawSpecies)) %>%
  replace_na(list(rawMoney = "", rawSpecies = "")) %>%
  unite("rawValue", rawMoney:rawSpecies, sep = "", remove = FALSE) %>%
  mutate(rawValue = as.numeric(rawValue),
         rawLabels = comma(rawValue))
fig.1.data_bio$rawLabels[c(seq(1,11,2))] <- str_glue("${round(rawFrame$rawMoney/1e6)}m")
# fig.1.data_bio$rawLabels[c(seq(2,12,2))] <- rawFrame$rawSpecies

fig.1.plot_bio <- ggplot(fig.1.data_bio, 
                      aes(fill=is.Money, 
                          y=Amount, 
                          x=Category)) + 
  geom_col(position="dodge") +
  labs(x="Threat Category", 
       y="Percentage") + 
       # title="% Money Allocated vs. % Species Threatened (IUCN) by Unique Threats") +
  scale_x_discrete(labels=u.threat_list) +
  scale_fill_manual(# values = c("#556B2F", "#8f3c36"),
                    values = alpha(c("darkgreen", "orangered4"), 0.9),
                    # values = c("#556B2F", "#708090"),
                    labels = c("Money ($USD Million)", "# Threatened Species")) +
  geom_text(position = position_dodge(width= 1), 
            aes(label = rawLabels, hjust = 0.5, vjust = -0.5), angle=0, size=6) +
  ylim(0, 50) +
  theme_minimal() +
  theme(axis.title = element_text(size = 16, face = "bold"), 
        axis.ticks.length.x = unit(.5, "cm"),
        axis.text = element_text(size = 14),
        legend.title = element_blank(),
        legend.key.size = unit(1,"cm"),
        legend.text = element_text(size = 12))

# ggsave(fig.1.plot_bio, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/money_species_comparison.png", width=15, height=7,dpi = 300)

fig.1.plot_bio
```


# Figure 2: Temporal Threat Analysis
## Money to Unique Threat Categories
### Line Graph

```{r, echo=FALSE}
threat_temporal.data <- by.year.money_bio %>%
  filter(Category %in% u.threat_cat)

threat_temporal.plot <- ggplot(threat_temporal.data, 
                aes(y = Amount, 
                    x = Year.Authorized,
                    color = Category)) +
  #geom_point() +
  geom_line(size = 2, linetype = 1) +
  # geom_smooth(size = 1, aes(fill = Category), show.legend = FALSE) +
  theme_minimal() +
  scale_color_manual(values = cat_colors22[c(1,13,11,20,5,18)],
                     name="Threat Category",
                     labels=u.threat_list) +
  scale_fill_manual(values = cat_colors7) +
  labs(x="Year", 
       y="Money ($US Million)", 
       title="Amount of Money Allocated Across Unique Threat Categories per Year") +
  scale_y_continuous(#labels = waiver(),
                     breaks = c(0,50e6,10e7),
                     labels = c("0","50","100"),
                     position = "right") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, face = "bold"),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.key.width = unit(1,"cm"),
        # legend.box.spacing = unit(1,"cm"),
        legend.spacing.x = unit(1,"cm"),
        legend.position = "bottom")
        # legend.background = element_rect(color = "black", size = 1.2, fill = NA))

# ggsave(threat_temporal.plot, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/threat_temporal.png", width=12, height=7,dpi = 300)


threat_temporal.plot
```

### Proportional (w smoothing option)
```{r, echo=FALSE}
threat_temporal.data <- by.year.money_bio %>%
  filter(Category %in% u.threat_cat) %>%
  group_by(Year.Authorized) %>%
  mutate(ann.percent = Amount/sum(Amount),
         alpha = rep(c(1,.5), 3))

threat_temporal.prop <- ggplot(threat_temporal.data, 
                aes(y = ann.percent, 
                    x = Year.Authorized)) +
                    # color = Category)) +
  # geom_point(aes(color = Category)) +
  # geom_line(size = 1, linetype = 5, alpha = 0.4, aes(color = Category)) +
  geom_line(size = 4, aes(color = Category, alpha = alpha), show.legend = F) +
  # geom_smooth(se = F, size = 2, linetype = 1, 
              # aes(color = Category, fill = Category)) +
  # theme_minimal() +
  scale_color_manual(values = cat_colors22[c(1,13,11,20,5,18)],
                     name="Threat Category",
                     labels=u.threat_list) +
  # scale_fill_manual(values = cat_colors22[c(1,13,11,20,5,18)],
  #                    name="Threat Category",
  #                    labels=u.threat_list) +
  # labs(x="Year", 
  #      y="Proportion of Funding", 
  #      title="Proportion of Funding") +
  theme(panel.background = element_rect(fill = "grey90"),
        # panel.border = element_rect(color = "black", fill = NA),
        # axis.line = element_line(color = "black"),
        axis.title = element_blank(),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 26))

# ggsave(threat_temporal.prop, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/threat_temporal_prop_inset.png", width=12, height=7,dpi = 300)


threat_temporal.prop
```

### Proportional change/year
```{r}
threat_clim <- threat_temporal.data %>%
       filter(Category == "cat_threat_climate") %>%
  ungroup() %>%
  mutate(per.change = (ann.percent - lag(ann.percent)))
print("Avg. Climate Change change/year")
mean(threat_clim$per.change, na.rm = T)

threat_hab <- threat_temporal.data %>%
  filter(Category =="cat_threat_habitat") %>%
  ungroup() %>%
  mutate(per.change = (ann.percent - lag(ann.percent)))
print("Avg. Habitat Loss change/year")
mean(threat_hab$per.change, na.rm = T)
```

# Figure 3: Taxa/Threat Maxwell Comparison
## Coercing to our species
```{r}
maxwell.s4 <- read_csv("~/Columbia E3B/Conservation Group/Maxwell et al table S4.csv")
names(maxwell.s4)

# Our species: Amphibians, Birds, Fish, Invertebrates, Mammals, Plants, Reptiles

our.species <- maxwell.s4 %>%
  slice(1:10) %>%
  mutate(Amphibians = Amphibia,
         Birds = Aves,
         Fish = Actinopterygii + Chondrichthyes,
         Invertebrates = Anthozoa + Gastropoda + Hydrozoa + Malacostraca + Myxini,
         Mammals = Mammalia,
         Plants = Cycadopsida + Liliopsida + Magnoliopsida + Pinopsida,
         Reptiles = Reptilia) %>%
  dplyr::select(c(2,3, 19:ncol(.), 18))
our.species
```

## Joint figure
```{r}
# Preparing maxwell data
fig.4.maxwell_data <- our.species %>% 
  dplyr::select(2:11) %>%
  group_by(Category) %>%
  summarise_at(vars(Total:Reptiles), sum) %>%
  dplyr::select(3:length(.)) %>%
  mutate(Total.maxwell = rowSums(.)) %>%
  dplyr::select(8,1:7) %>%
  transmute_at(vars(-Total.maxwell), ~ (./Total.maxwell)*100) %>%
  mutate(Threat = u.threat_list) %>%
  dplyr::select(8,1:7) %>%
  pivot_longer(Amphibians:Reptiles, names_to = "Taxa", values_to = "Percent") %>%
  mutate(Source = "IUCN", direction = -1, alpha = 0.9)


# Preparing money data
taxa_climate <- money.data_bio %>%
  filter(cat_threat_climate > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_climate") %>%
  `colnames<-` (u.taxa_list)

taxa_energy <- money.data_bio %>%
  filter(cat_threat_energytransport > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_energytransport") %>%
  `colnames<-` (u.taxa_list)

taxa_habitat <- money.data_bio %>%
  filter(cat_threat_habitat > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_habitat") %>%
  `colnames<-` (u.taxa_list)

taxa_inv <- money.data_bio %>%
  filter(cat_threat_inv_spec > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_inv_spec") %>%
  `colnames<-` (u.taxa_list)

taxa_overexp <- money.data_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_overexp") %>%
  `colnames<-` (u.taxa_list)

taxa_pollution <- money.data_bio %>%
  filter(cat_threat_pollution > 0) %>%
  dplyr::select(cat_taxa_amphibian:cat_taxa_reptile) %>%
  pivot_longer(cat_taxa_amphibian:cat_taxa_reptile, 
               names_to = "Category", values_to = "Money") %>%
  group_by(Category) %>%
  summarize(totalAmount = sum(Money)) %>%
  mutate(perc = totalAmount/sum(totalAmount)) %>%
  dplyr::select(perc) %>%
  t() %>%
  as.data.frame(row.names = "cat_threat_pollution") %>%
  `colnames<-` (u.taxa_list)

fig.4.money_data <- bind_rows(taxa_climate, taxa_energy, taxa_habitat, 
                         taxa_inv, taxa_overexp, taxa_pollution) %>%
  transmute_all(~.*100) %>%
  mutate(Threat = u.threat_list) %>%
  dplyr::select(8, 1:7) %>%
  pivot_longer(Amphibians:Reptiles, names_to = "Taxa", values_to = "Percent") %>%
  mutate(Source = "Money", direction = 1, alpha = 1)

### 1 plot
fig.4.data <- bind_rows(fig.4.money_data, fig.4.maxwell_data) %>%
  mutate(Percent = Percent * direction,
         Taxa = factor(Taxa, 
                       levels = c("Reptiles", "Amphibians", "Invertebrates", 
                                  "Birds", "Mammals", "Fish", "Plants"),
                       ordered = TRUE),
         Threat = factor(Threat,
                         levels = c("Climate Change", "Overexploitation",
                                    "Habitat Loss", "Energy & Transportation",
                                    "Pollution", "Invasive Species"),
                         ordered = TRUE)
  )

fig.4 <- ggplot(fig.4.data %>% filter(Taxa != "Reptiles"),
                aes(x = Taxa, y = Percent, fill = Taxa)) + 
  geom_col(show.legend = F , color = "black") +
  geom_hline(yintercept = 0, size = 1.2) +
  scale_fill_manual(values = cat_colors22[c(9,20,19,15,7,11)],
                    name = "Taxa",
                    labels = u.taxa_list) +
  facet_wrap(~Threat) +
  theme_bw() +
  theme(axis.ticks.y= element_blank(),
        # axis.text.y = element_text(size = 10, face = "bold"),
        axis.text.x = element_blank(),
        # axis.text.y = element_text(size = 12),
        # axis.title = element_text(size = 16, face = "bold"),
        axis.ticks.x=element_line(color=NA),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(size = 12, color = "white", face = "bold")) +
  ylim(-90,90) +
  ggtitle("Percent of Taxa Threatened (IUCN) by vs. Funded for Unique Threats") +
  ylab("% Taxa Threatened   |    % Funding to Taxa") +
  coord_flip()

# ggsave(fig.4, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/taxa_threat_comparison.png", width=12, height=7,dpi = 300)

fig.4

```

# Figure 4: Threats Network Analysis
## Threat Category Overlap
```{r, include=FALSE}
threat.net1 <- data.multi_bio %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002) %>%
  dplyr::select(cat_threat_climate:cat_threat_pollution)
threat.net1 <- map2_df(threat.net1, colnames(threat.net1), ~  c(NA, .y)[.x +1]) # Change the binary 1 values into the column heaers
threat.net2 <- data.multi_bio %>%
  filter(Year.Authorized < 2017) %>%
  filter(Year.Authorized != 2002) %>%
  dplyr::select(Grant.Amount, Num.Thr)
threat.net3 <- bind_cols(threat.net2, threat.net1) # Join the column-header data with the grant amount/number of threats per grant

threat.net3 <- threat.net3 %>%
  filter(Num.Thr != 0 & Num.Thr <= 2) %>%
  unite("z", sep = ",", cat_threat_climate:cat_threat_pollution, na.rm = TRUE, remove = FALSE) %>% # Bring together all columns into a single column, separate contents by a comma
  dplyr::select(Grant.Amount, Num.Thr, z) %>%
  separate(z, sep = ",", c("A", "B")) %>% # Break united column into two
  mutate(B = if_else(is.na(B), A, B)) %>% # Have all NA's = Column A
  mutate(A = case_when(
      A == "cat_threat_climate" ~ "Climate Change",
      A == "cat_threat_energytransport" ~ "Energy & Transportation",
      A == "cat_threat_habitat" ~ "Habitat Loss",
      A == "cat_threat_inv_spec" ~ "Invasive Species",
      A == "cat_threat_overexp" ~ "Overexploitation",
      A == "cat_threat_pollution" ~ "Pollution")) %>%
  mutate(B = case_when(
      B == "cat_threat_climate" ~ "Climate Change",
      B == "cat_threat_energytransport" ~ "Energy & Transportation",
      B == "cat_threat_habitat" ~ "Habitat Loss",
      B == "cat_threat_inv_spec" ~ "Invasive Species",
      B == "cat_threat_overexp" ~ "Overexploitation",
      B == "cat_threat_pollution" ~ "Pollution")) %>%
  group_by(A, B) %>% # group by the combinations between column A and column B, which contain one or two threats
  summarize(n = sum(Grant.Amount)) %>%
  filter(A != B) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

# write_csv(threat.net3, "~/Columbia E3B/Conservation Group/Output V15.2/Tables/threat_overlap_bio.csv")

threat.net.fig <- threat.net3 %>%
  ggplot(aes(x = A, y = B)) +
  geom_tile(aes(fill = log10(n)), color = "black") +
  geom_text(aes(label = n, fontface = "bold"), color = "white") +
  scale_fill_gradient(low = "lightgray", high = "black") +
  theme_bw() +
  labs(x="",
       y = "",
       title = "Amount of Money ($USD) in Single and Paired Threat Categories") +
  theme(legend.position = "none",
        panel.grid.major = element_blank())

threat.net.fig
```

## Pie Charts
```{r}
blank_theme <- theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        plot.title=element_text(size=14, face="bold"))
```

### # of Threats within Threat categories (number of grants, not $)
#### Climate
```{r}
pie.Nthreats_climate <- money.data.multi_bio %>%
  filter(cat_threat_climate > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount),
         alpha = seq(0.2,1,0.2))

plot.pie.Nthreats_climate <- ggplot(pie.Nthreats_climate,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[1], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_manual(values = c("#a50f15","#de2d26","#fb6a4a","#fcae91","#fee5d9")) +
  # scale_fill_viridis_d(name = "Number of Threats", option = "D") +
  labs(title="Proportion of Climate Change Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

# ggsave(plot.pie.Nthreats_climate, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_climate.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_climate
```


#### Energy/Transportation
```{r}
pie.Nthreats_energytransport <- money.data.multi_bio %>%
  filter(cat_threat_energytransport > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_energytransport <- ggplot(pie.Nthreats_energytransport,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "black") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[13], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Energy/Transportation Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

# ggsave(plot.pie.Nthreats_energytransport, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_energytransport.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_energytransport
```


#### Habitat Loss
```{r}
pie.Nthreats_habitat <- money.data.multi_bio %>%
  filter(cat_threat_habitat > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_habitat <- ggplot(pie.Nthreats_habitat,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
    scale_fill_manual(values = seq_gradient_pal(cat_colors22[11], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats", end = 0.8) +
  labs(title="Proportion of Habitat Loss Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4)),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_habitat, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_habitat.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_habitat
```


#### Invasive Species
```{r}
pie.Nthreats_inv_spec <- money.data.multi_bio %>%
  filter(cat_threat_inv_spec > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_inv_spec <- ggplot(pie.Nthreats_inv_spec,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "black") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[20], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Invasive Species Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_inv_spec, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_inv_spec.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_inv_spec
```

#### Overexploitation
```{r}
pie.Nthreats_overexp <- money.data.multi_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_overexp <- ggplot(pie.Nthreats_overexp,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[5], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Overexploitation Grants Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                # label = paste0(round(perc*100,1), "%")),
                label = str_glue("{round(perc*100,1)}% ({totalCount})")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_overexp, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_overexp.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_overexp
```

##### Prop Overxploitation funding
```{r}
pie.threats_overexp <- money.data.multi_bio %>%
  filter(cat_threat_overexp > 0) %>%
  dplyr::select(Num.Thr, cat_threat_overexp) %>%
  group_by(Num.Thr) %>%
  summarize(total = sum(cat_threat_overexp)) %>%
  mutate(perc = total/sum(total))
```


#### Pollution
```{r}
pie.Nthreats_pollution <- money.data.multi_bio %>%
  filter(cat_threat_pollution > 0) %>%
  dplyr::select(Num.Thr) %>%
  mutate(Num.Thr = as.factor(Num.Thr)) %>%
  mutate(Count = 1) %>%
  group_by(Num.Thr) %>%
  summarize(totalCount = sum(Count)) %>%
  mutate(perc = totalCount/sum(totalCount))

plot.pie.Nthreats_pollution <- ggplot(pie.Nthreats_pollution,
                       aes(y = perc, 
                           x = "",
                           fill = Num.Thr)) +
  geom_bar(stat = "identity", color = "white") +
  theme_minimal() +
  scale_fill_manual(values = seq_gradient_pal(cat_colors22[18], "white",
                                              "Lab")(seq(0,1, length.out = 5))) +
  # scale_fill_viridis_d(name = "Number of Threats") +
  labs(title="Proportion of Pollution Funding Classified into Multiple Threats") +
  coord_polar("y", start=0, direction = -1) +
  blank_theme +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(x = c(rep(1.55,4), 1.65),
                label = paste0(round(perc*100,1), "%")),
            position = position_stack(vjust = 0.5),
            size = 3,
            col = "black")

ggsave(plot.pie.Nthreats_pollution, filename = "~/Columbia E3B/Conservation Group/Output V15.2/Figures/pie_Nthreats_pollution.png", width=12, height=7,dpi = 300)

plot.pie.Nthreats_pollution
```

# Figure S1
```{r}
donor_Rank <- c(" ", "Top 50", "Top 100", "Top 200", "Top 300", "Top 400", "Top 500", "Top 600", "Top 700", "Top 800", "Top 900", "Top 1000")
sum <- c(0, 14396513202, 16440571398, 18171144637, 18975986509, 19490663536, 19851261916, 20128241008, 20348022885, 20525790876, 20670921174, 20793224389)
sum_mildiv <- sum/1000000
axis_Order <- c(1:12)

sum.data <- data.frame(donor_Rank, sum, sum_mildiv, axis_Order)

sum_graph <- ggplot(data = sum.data,
                    mapping = aes(x = axis_Order,
                                  y =  sum_mildiv,
                                  group = 1)
) +
  geom_line() +
  geom_point() +
  geom_area(aes(y=ifelse(sum_mildiv>=0 & sum_mildiv< 19851.26, sum_mildiv, 19851.26)), fill="#696969", alpha = .4) +
  ylim(0, 21500) +
  theme_minimal() +
  ggtitle("Total Biodiversity/Climate Change Grantmaking by Top 1000 Grantmakers") +
  xlab("Grantmaker Group") +
  ylab("Sum ($1,000,000USD)") + 
  scale_x_discrete(limits=c(" ", "Top 50", "Top 100", "Top 200", "Top 300", "Top 400", "Top 500", "Top 600", 
                                      "Top 700", "Top 800", "Top 900", "Top 1000"))

sum_graph
```


# Miscellaneous Analyses
## Overexploitation Analysis
```{r}
overexp <- data_bio %>%
  filter(cat_threat_overexp ==1)

fishery <- data_bio %>%
  filter(cat_threat_overexp ==1) %>%
  filter(str_detect(Description, "fishery|fisheries"))

print("fishery prop")
sum(fishery$Grant.Amount)/sum(data_bio %>% filter(cat_threat_overexp ==1)%>%dplyr::select(Grant.Amount))

industry <- data_bio %>%
  filter(cat_threat_overexp ==1) %>%
  filter(str_detect(Description, "fishery|fisheries|forestry"))

print("industry prop")
sum(industry$Grant.Amount)/sum(data_bio %>% filter(cat_threat_overexp ==1)%>%dplyr::select(Grant.Amount))
print("industry median")
median(industry$Grant.Amount)
print("industry mean")
mean(industry$Grant.Amount)
print("overexploitation median")
median(overexp$Grant.Amount)
print("overexploitation mean")
mean(overexp$Grant.Amount)

```

## Threat Category Analysis
```{r}
multi_threat <- data_bio %>% filter(Num.Thr >1)

print("Prop of Threat Money")
sum(multi_threat$Grant.Amount)/
  sum(data_bio %>% filter(Num.Thr > 0) %>% dplyr::select(Grant.Amount))

print("Prop of Total Money")
sum(multi_threat$Grant.Amount)/
  sum(data_bio %>% dplyr::select(Grant.Amount))
```

## Climate Change Analysis
```{r}
climate.sub.1 <- data_bio %>% 
  filter(cat_threat_climate ==1) %>% 
  filter(str_detect(Description, "land management|restoration|natural climate solutions|biodiversity conservation"))

climate.sub.1.multi <- data.multi_bio %>% filter(cat_threat_climate ==1) %>% filter(str_detect(Description, "land management|restoration|natural climate solutions|biodiversity conservation"))
```

## Multicategory Analysis
```{r}
clim_hab1 <- data_bio %>%
  filter(cat_threat_climate_habitat ==1) %>%
  dplyr::select(c(1:10, 104:109, 67))
clim_hab2 <- data.multi_bio %>%
  filter(cat_threat_climate ==1 & cat_threat_habitat ==1 & Num.Thr ==2) %>%
  dplyr::select(c(1:10, 43:48)) %>%
  mutate(cat_threat_climate_habitat =1)

difference <- anti_join(clim_hab1, clim_hab2)
```

## Check data/data_multi
```{r}
data_bio_test <- data_bio %>%
  dplyr::select(c(1:10, 104:109))
data.multi_bio_test <- data.multi_bio %>%
  dplyr::select(c(1:10, 43:48))

multi_diff <- anti_join(data.multi_bio_test, data_bio_test)

data_test <- data %>%
  dplyr::select(c(1:10, 104:109))
data.multi_test <- data.multi %>%
  dplyr::select(c(1:10, 43:48))

multi_diff2 <- anti_join(data.multi_test, data_test)

### Only 1 grant in data.multi that isn't in data (threat_other but not CC/Energy)
```